---
title: "Flux Calculation for real time gas analyzers"
author: "Holly Andrews and Alex Krichels"
date: "December, 2020"
output: pdf_document
editor_options: 
  chunk_output_type: console
---
Goal:  Import soil climate, CO2, N2O, and NOx raw data along with LI-8100 chamber measurement times (hereafter, "summary" data). 


```{r packages}
library(chron)
library(dplyr)
library(lubridate)
library(ggplot2)
library(cowplot)
library(DescTools)
```

Make a function that loops through CO2 and N2O data and extracts rows for a 90-second interval (90 rows total) of N2O data.  UPDATE: This function now calculates flux for each 90-second interval and prints the slope (flux), intercept, and R2 in the new dataset!  Makes calculations easier to see quickly.

```{r functions}
#CO2 flux
extractCO2 <- function(smalldata, bigdata) { #smalldata = chamber measurement times (summary data), bigdata = raw 8100 values
  df <- data.frame()
  for (row in 1:nrow(smalldata)) {
    endrow <- smalldata$DateNum[row]
    beginrow <- endrow-90 #need to hardcode whatever interval to calculate flux over, may depend on experimental conditions; this is the number of seconds in your measurement window
    end <- bigdata[which.min(abs(bigdata$DateNum - endrow)),]
    begin <- bigdata[which.min(abs(bigdata$DateNum-beginrow)),]
    sub <- bigdata[which(bigdata$DateNum == begin$DateNum):which(bigdata$DateNum == end$DateNum),]
    sub$TSB <- sub$DateNum-sub$DateNum[1] #TSB = Time Since Beginning (of measurement)
    if(count(sub)>5) { #ensures that you have at least 5 data points in 90-second measurement time 
       temp <- mean(sub$Temp)
       moist <- mean(sub$H2O)
       chamTemp <- mean(sub$Tcham)
       pCO2 <- summary(lm(sub$Cdry~sub$TSB))
       mCO2 <- pCO2$coefficients[2]
       bCO2 <- pCO2$coefficients[1]
       rCO2 <- pCO2$r.squared
       #if(sub$Cdry[1]<500 | mCO2>0) {  #has only been used in one dataset of Holly's data
       lines <- data.frame(smalldata[row,],mCO2,bCO2,temp,moist,rCO2,chamTemp)
       df <- rbind(df,lines)
       #}
    }
  } 
  return(df)
}

#N2O flux
extractN2O <- function(smalldata, bigdata) {
  df <- data.frame()
  for (row in 1:nrow(smalldata)) {
    endrow <- smalldata$DateNum[row]-10
    beginrow <- endrow-90 #need to hardcode whatever interval to calculate flux over, may depend on experimental conditions
    end <- bigdata[which.min(abs(bigdata$DateNum - endrow)),]
    begin <- bigdata[which.min(abs(bigdata$DateNum-beginrow)),]
    sub <- bigdata[which(bigdata$DateNum == begin$DateNum):which(bigdata$DateNum == end$DateNum),]
    sub$TSB <- sub$DateNum-sub$DateNum[1]
    if(count(sub)>5) {
       pN2O <- summary(lm(sub$N2O~sub$TSB))
       mN2O <- pN2O$coefficients[2]
       bN2O <- pN2O$coefficients[1]
       rN2O <- pN2O$r.squared
       lines <- data.frame(smalldata[row,],mN2O,bN2O,rN2O)
       df <- rbind(df,lines)
    }
  } 
  return(df)
}

#N2O iso - this is taking the average of the last 10 seconds for the isotope measurements
extractISO <- function(smalldata, bigdata) {
  df <- data.frame()
  for (row in 1:nrow(smalldata)) {
    endrow <- smalldata$DateNum[row]-10
    beginrow <- endrow-10 #need to hardcode whatever interval to calculate flux over, may depend on experimental conditions
    end <- bigdata[which.min(abs(bigdata$DateNum - endrow)),]
    begin <- bigdata[which.min(abs(bigdata$DateNum-beginrow)),]
    sub <- bigdata[which(bigdata$DateNum == begin$DateNum):which(bigdata$DateNum == end$DateNum),]
    sub$TSB <- sub$DateNum-sub$DateNum[1]
    if(count(sub)>1) {
       d15N <- mean(sub$d15N)
       lines <- data.frame(smalldata[row,],d15N)
       df <- rbind(df,lines)
    }
  } 
  return(df)
}

#NOx flux
extractNOx <- function(smalldata, bigdata) {
  df <- data.frame()
  for (row in 1:nrow(smalldata)) {
    endrow <- smalldata$DateNum[row]
    beginrow <- endrow-90 #need to hardcode whatever interval to calculate flux over, may depend on experimental conditions
    end <- bigdata[which.min(abs(bigdata$DateNum - endrow)),]
    begin <- bigdata[which.min(abs(bigdata$DateNum-beginrow)),]
    sub <- bigdata[which(bigdata$DateNum == begin$DateNum):which(bigdata$DateNum == end$DateNum),]
    sub$TSB <- sub$DateNum-sub$DateNum[1]
    if(count(sub)>4) { #4 data point minimum since NOx data is in 10-second intervals
      pNO <- summary(lm(sub$NO~sub$TSB))
      mNO <- pNO$coefficients[2]
      bNO <- pNO$coefficients[1]
      pNO2 <- summary(lm(sub$NO2~sub$TSB))
      mNO2 <- pNO2$coefficients[2]
      bNO2 <- pNO2$coefficients[1]
      rNO <- pNO$r.squared
      rNO2 <- pNO2$r.squared
      lines <- data.frame(smalldata[row,],mNO,bNO,mNO2,bNO2,rNO,rNO2)
      df <- rbind(df,lines)
    }
  } 
  return(df)
}
  
extrapolateFlux <- function(data, offset) {
  C_g_mol <- 12
  N_g_mol <- 14.0067
  R_constant <- 0.0821
  chamber_area <- 0.03178 
  volume <- (19.0 + 4076.2 + 55.0 + 237.0 + (324.13*offset))/1000
  data$CO2flux <- data$mCO2 * volume * C_g_mol / chamber_area / R_constant / (data$chamTemp + 273.15) #Final units: ug CO2-C m^-2 s^-1
  data$N2Oflux <- data$mN2O * volume * (2*N_g_mol) / chamber_area / R_constant /(data$chamTemp + 273.15)#Final units: ug N2O-N m^-2 s^-1
  data$NOflux <- data$mNO * volume * N_g_mol / chamber_area / R_constant / (data$chamTemp + 273.15) #Final units: ng NO-N m^-2 s^-1
  return(data)
}

#Volume includes all instrumets etc. in cubic cm

```


PART I: RAW DATA COMPILATION AND STORAGE


This section imports individual CO2, N2O, and NOx files and merges them into larger files which are exported.  Once files are written, this section shouldn't need to be run unless changes are made to raw data.

```{r import-CO2-subset}
setwd("C:/Users/ahkri/Licor_Data")

#raw 8100 data
CO2datafile <- read.delim("2018_Licor_NSpiral_raw.txt",sep="\t",header=TRUE,skip=34) #number of skip lines may need to be amended depending on dataset, usually skip is 29 or 30 lines; if error generated, try different # skip lines

CO2datanohead <- subset(CO2datafile,Type=="1") #gets rid of the header, footer, and summary

CO2datasub <- data.frame(CO2datanohead$Date,CO2datanohead$Cdry,CO2datanohead$V2,CO2datanohead$V3, CO2datanohead$Tcham)
colnames(CO2datasub) <- c("Date","Cdry","Temp","H2O","Tcham")
CO2datasub <- arrange(CO2datasub,Date)
CO2datasub$Date <- as.POSIXct(CO2datasub$Date, format="%Y-%m-%d %H:%M:%S")
CO2datasub <- subset(CO2datasub, Date > as.POSIXct("2018-08-21 08:04:40"))
CO2datasub <- subset(CO2datasub, Date < as.POSIXct("2018-08-22 08:30:01"))


#CO2 summary data
CO2sum <- read.delim("2018_Licor_NSpiral_sum.txt",sep="\t",header=TRUE)
CO2sumsub <- data.frame(CO2sum$Port,CO2sum$Date_IV,CO2sum$DOY_IV,CO2sum$Exp_Flux, CO2sum$Exp_Flux)
colnames(CO2sumsub) <- c("Port.","Date","DOY","Exp_flux","Exp_flux_1")
CO2sumsub$Date <- as.POSIXct(CO2sumsub$Date, format="%Y-%m-%d %H:%M:%S")
CO2sumsub <- subset(CO2sumsub, Date > as.POSIXct("2018-08-21 08:04:40")) #This time frame works with the example dataset
CO2sumsub <- subset(CO2sumsub, Date < as.POSIXct("2018-08-22 08:30:01"))


write.table(CO2datasub, file = "C:/Users/ahkri/R_Outputs/CO2raw.txt", sep = ",", row.names = FALSE) # "bigdata"

write.table(CO2sumsub, file = "C:/Users/ahkri/R_Outputs/CO2sum.txt", sep = ",", row.names = FALSE)  # "Smalldata"
```

```{r import-N2O-subset}
#raw N2O data
N2Odatafolder <- "C:/Users/ahkri/LGR_Data/"
N2Olist <- list.files(path=N2Odatafolder, pattern="*.txt")
N2Odata <- do.call("rbind", lapply(N2Olist, function(x) read.delim(paste(N2Odatafolder, x, sep=''),sep=",",skip=1,header=TRUE,stringsAsFactors = FALSE)))
N2Odata2 <- subset(N2Odata, X.N2O._ppm > 0) #subsetting to remove footer and other useless rows

write.table(N2Odata2, file = "C:/Users/ahkri/OneDrive/Desktop/Projects/POSTDOC/N SPiraling/2018_Fluxes/Boyd/Nitrate/R_Outputs/N2Oraw.txt", sep = ",", row.names = FALSE) #

N2Odatasub <- data.frame(N2Odata2$Time, N2Odata2$X.N2O._ppm, N2Odata2$d15N)
colnames(N2Odatasub) <- c("Date","N2O", "d15N")

write.table(N2Odatasub, file = "C:/Users/ahkri/R_Outputs/N2Oraw2col.txt", sep = ",", row.names = FALSE) #"bigdata" for N2O
```

```{r import-NOx-subset}
#raw NOx data 
NOxdata <- read.delim("C:/Users/ahkri/NOx_Data/NoxBox_Data5sec.dat",skip=1,header=TRUE,sep=",")
NOxdata$TIMESTAMP <- as.POSIXct(NOxdata$TIMESTAMP, format="%Y-%m-%d %H:%M:%S")
NOxsub <- subset(NOxdata, TIMESTAMP > as.POSIXct("2018-08-20 00:00:00")) # since this file is huge, I subsetted only to campaign-specific data
NOxsub2 <- data.frame(NOxsub$TIMESTAMP,NOxsub$NO2_Avg,NOxsub$NO_Avg,NOxsub$NOx_Avg)
colnames(NOxsub2) <- c("Date","NO2","NO","NOx")

write.table(NOxsub2, file = "C:/Users/ahkri/R_Outputs/NOxraw.txt", sep = ",", row.names = FALSE)

```

To save memory in R, suggest to clear environment at this point.


PART II: CHAMBER FLUX CALCULATIONS
Re-import raw and summary data and use extract functions to calculate individual chamber flux measurements (i.e. regression slopes).  Then add to new dataframe.

```{r import-extract}
setwd("C:/Users/ahkri/R_Outputs/")
CO2data <- read.delim("CO2raw.txt",header=TRUE,sep=",")
CO2sum <- read.delim("CO2sum.txt",header=TRUE,sep=",")
N2Odata <- read.delim("N2Oraw2col.txt",header=TRUE,sep=",")
NOxdata <- read.delim("NOxraw.txt",header=TRUE,sep=",")

CO2data$Date <- as.POSIXct(CO2data$Date,format= "%Y-%m-%d %H:%M:%S")
CO2sum$Date <- as.POSIXct(CO2sum$Date)
N2Odata$Date <- as.POSIXct(N2Odata$Date,format=" %m/%d/%Y %H:%M:%S")
NOxdata$Date <- as.POSIXct(NOxdata$Date)

CO2data$DateNum <- as.numeric(CO2data$Date)
CO2sum$DateNum <- as.numeric(CO2sum$Date)
N2Odata$DateNum <- as.numeric(N2Odata$Date)
NOxdata$DateNum <- as.numeric(NOxdata$Date)

CO2data$Cdry <- as.numeric(as.character(CO2data$Cdry))
CO2data$Temp <- as.numeric(as.character(CO2data$Temp))
CO2data$H2O <- as.numeric(as.character(CO2data$H2O))
N2Odata$N2O <- as.numeric(as.character(N2Odata$N2O))
N2Odata$d15NA <- as.numeric(as.character(N2Odata$d15NA))
N2Odata$d15NB <- as.numeric(as.character(N2Odata$d15NB))
N2Odata$d15N <- as.numeric(as.character(N2Odata$d15N))
N2Odata$SP <- as.numeric(as.character(N2Odata$SP))
N2Odata$d18O <- as.numeric(as.character(N2Odata$d18O))
N2Odata$d17O <- as.numeric(as.character(N2Odata$d17O))


NOxdata$NO2 <- as.numeric(as.character(NOxdata$NO2))
NOxdata$NO <- as.numeric(as.character(NOxdata$NO))
NOxdata$NOx <- as.numeric(as.character(NOxdata$NOx))
#NOxdata2 <- subset(NOxdata, NO2 > -200 & NO2 < 1000000) #if NOx data needs cleaning; this data can be problematic
#NOxdatanona <- NOxdata2[complete.cases(NOxdata2),] # subsetted to just cases where all cells in a row are filled


###

#If instrument times are not perfectly synced then they may need to be corrected. 

N2Odata$DateNum <- (N2Odata$DateNum + 140) #insert offset here (in seconds) if it is a fixed time. This offset works for the example dataset
NOxdata$DateNum <- (NOxdata$DateNum  + 160) #insert offset here (in seconds) if it is a fixed time. This offset works for the example dataset
###

#subset summary data by ports so that it is easier to add treatment, port, etc., info before merging
port1 <- subset(CO2sum, Port.=="1")
port2 <- subset(CO2sum, Port.=="2")
port3 <- subset(CO2sum, Port.=="3")
port4 <- subset(CO2sum, Port.=="4")
port5 <- subset(CO2sum, Port.=="5")
port6 <- subset(CO2sum, Port.=="6")
port7 <- subset(CO2sum, Port.=="7")
port8 <- subset(CO2sum, Port.=="8")

#run extraction functions for each port
CO2ex1 <- extractCO2(port1,CO2data)
N2Oex1 <- extractN2O(port1,N2Odata)
N2OISO1 <- extractISO(port1,N2Odata)
NOxex1 <- extractNOx(port1,NOxdata)
port1CO2N2O <- merge(CO2ex1,N2Oex1, by="DateNum",all=TRUE) # the "all" argument keeps all data even if it exists for CO2 and not N2O or vice versa
port1CO2N2OISO <- merge(port1CO2N2O, N2OISO1, by="DateNum",all=TRUE)

port1gas <- merge(port1CO2N2OISO,NOxex1,by="DateNum",all=TRUE)

CO2ex2 <- extractCO2(port2,CO2data)
N2Oex2 <- extractN2O(port2,N2Odata)
N2OISO2 <- extractISO(port2,N2Odata)
NOxex2 <- extractNOx(port2,NOxdata)
port2CO2N2O <- merge(CO2ex2,N2Oex2, by="DateNum",all=TRUE)
port2CO2N2OISO <- merge(port2CO2N2O, N2OISO2, by="DateNum",all=TRUE)
port2gas <- merge(port2CO2N2OISO,NOxex2,by="DateNum",all=TRUE)

CO2ex3 <- extractCO2(port3,CO2data)
N2Oex3 <- extractN2O(port3,N2Odata)
N2OISO3 <- extractISO(port3,N2Odata)
NOxex3 <- extractNOx(port3,NOxdata)
port3CO2N2O <- merge(CO2ex3,N2Oex3, by="DateNum",all=TRUE)
port3CO2N2OISO <- merge(port3CO2N2O, N2OISO3, by="DateNum",all=TRUE)
port3gas <- merge(port3CO2N2OISO,NOxex3,by="DateNum",all=TRUE)

CO2ex4 <- extractCO2(port4,CO2data)
N2Oex4 <- extractN2O(port4,N2Odata)
N2OISO4 <- extractISO(port4,N2Odata)
NOxex4 <- extractNOx(port4,NOxdata)
port4CO2N2O <- merge(CO2ex4,N2Oex4, by="DateNum",all=TRUE)
port4CO2N2OISO <- merge(port4CO2N2O, N2OISO4, by="DateNum",all=TRUE)
port4gas <- merge(port4CO2N2OISO,NOxex4,by="DateNum",all=TRUE)

CO2ex5 <- extractCO2(port5,CO2data)
N2Oex5 <- extractN2O(port5,N2Odata)
N2OISO5 <- extractISO(port5,N2Odata)
NOxex5 <- extractNOx(port5,NOxdata)
port5CO2N2O <- merge(CO2ex5,N2Oex5, by="DateNum",all=TRUE)
port5CO2N2OISO <- merge(port5CO2N2O, N2OISO5, by="DateNum",all=TRUE)
port5gas <- merge(port5CO2N2OISO,NOxex5,by="DateNum",all=TRUE)

CO2ex6 <- extractCO2(port6,CO2data)
N2Oex6 <- extractN2O(port6,N2Odata)
N2OISO6 <- extractISO(port6,N2Odata)
NOxex6 <- extractNOx(port6,NOxdata)
port6CO2N2O <- merge(CO2ex6,N2Oex6, by="DateNum",all=TRUE)
port6CO2N2OISO <- merge(port6CO2N2O, N2OISO6, by="DateNum",all=TRUE)
port6gas <- merge(port6CO2N2OISO,NOxex6,by="DateNum",all=TRUE)

CO2ex7 <- extractCO2(port7,CO2data)
N2Oex7 <- extractN2O(port7,N2Odata)
N2OISO7 <- extractISO(port7,N2Odata)
NOxex7 <- extractNOx(port7,NOxdata)
port7CO2N2O <- merge(CO2ex7,N2Oex7, by="DateNum",all=TRUE)
port7CO2N2OISO <- merge(port7CO2N2O, N2OISO7, by="DateNum",all=TRUE)
port7gas <- merge(port7CO2N2OISO,NOxex7,by="DateNum",all=TRUE)

CO2ex8 <- extractCO2(port8,CO2data)
N2Oex8 <- extractN2O(port8,N2Odata)
N2OISO8 <- extractISO(port8,N2Odata)
NOxex8 <- extractNOx(port8,NOxdata)
port8CO2N2O <- merge(CO2ex8,N2Oex8, by="DateNum",all=TRUE)
port8CO2N2OISO <- merge(port8CO2N2O, N2OISO8, by="DateNum",all=TRUE)
port8gas <- merge(port8CO2N2OISO,NOxex8,by="DateNum",all=TRUE)

#add port and other information
    port1gas$Port. <- "1"
    port2gas$Port. <- "2"
    port3gas$Port. <- "3"
    port4gas$Port. <- "4"
    port5gas$Port. <- "5"
    port6gas$Port. <- "6"
    port7gas$Port. <- "7"
    port8gas$Port. <- "8"

#these numbers manually added to show when water was added to soil    
        port1gas$TimeSinceWetting <- (port1gas$DateNum-1534866071)
        port2gas$TimeSinceWetting <- (port2gas$DateNum-1534866313)
        port3gas$TimeSinceWetting <- (port3gas$DateNum-1534866556)
        port4gas$TimeSinceWetting <- (port4gas$DateNum-1534866798)
        port5gas$TimeSinceWetting <- (port5gas$DateNum-1534867039)
        port6gas$TimeSinceWetting <- (port6gas$DateNum-1534867282)
        port7gas$TimeSinceWetting <- (port7gas$DateNum-1534867525)
        port8gas$TimeSinceWetting <- (port8gas$DateNum-1534867766)

    
gasslopes <- rbind(port1gas,port2gas,port3gas,port4gas,port5gas,port6gas,port7gas,port8gas)

write.table(gasslopes, file = "C:/Users/ahkri/Routputs/chamberfluxes.txt", sep = ",", row.names = FALSE)

```

PART III: FLUX EXTRAPOLATION.   Here we need to subset by port to incorporate collar offsets; the best way to do this is probably to calculate each port's fluxes separately and then merge all together again.  To calculate Vtotals, go into header info within a random 8100 observation within file viewer and recalculate based off of chamber offsets.

```{r full-flux}
setwd("C:/Users/ahkri/R_outputs")
flux <- read.delim("chamberfluxes.txt",header=TRUE,sep=",")

port1 <- subset(flux, Port.=="1")
port2 <- subset(flux, Port.=="2")
port3 <- subset(flux, Port.=="3")
port4 <- subset(flux, Port.=="4")
port5 <- subset(flux, Port.=="5")
port6 <- subset(flux, Port.=="6")
port7 <- subset(flux, Port.=="7")
port8 <- subset(flux, Port.=="8")


flux1 <- extrapolateFlux(port1, 6.00)   #note, 6 is an example here - update these numbers based on chamber and collar volume
flux2 <- extrapolateFlux(port2, 6.00)
flux3 <- extrapolateFlux(port3, 6.00)
flux4 <- extrapolateFlux(port4, 6.00)
flux5 <- extrapolateFlux(port5, 6.00)
flux6 <- extrapolateFlux(port6, 6.00)
flux7 <- extrapolateFlux(port7, 6.00)
flux8 <- extrapolateFlux(port8, 6.00)

# Beolow is an example of how to calculate the  chamber volume, which is inclued above as the second bumber in the extrapolate flux function. 
         # CollarConstant = (3.14*(10.16^2))  # this is just PI * r2 part of volume calculation for 8in collar 
           # C1O <- mean(c(1.3,2.3,2.5))  #enter inner offsets here
           # C1I <- mean(c(5.5,8.3,9.1)) # enter outer offsets here
          #  C1Vol <- (CollarConstant*(C1I-C1O)/1000)


fluxall <- rbind(flux1,flux2,flux3,flux4,flux5,flux6,flux7,flux8)

write.table(fluxall, file = "C:/Users/ahkri/R_outputs/fluxes.txt", sep = ",", row.names = FALSE)

```

NEXT STEPS: Visualization and analysis. See other .Rmd files for code on plotting.



































